---
title: "LCA_PE"
author: "Vasileios Stavropoulos & Taylor Brown"
date: "2023-08-23"
output: html_document
---
#Import Library and Data
```{r}
message=FALSE
echo=TRUE
warnings=FALSE
error=FALSE

library(haven) ## Reading data - importing datasets
library(tidyLPA) ## Run LPA
library(ggplot2) ## Plotting
library(tidyverse) ## Enables tidy principles
library(mice) ## Missing data and imputations
library(naniar) ## MCAR test
library(stats)

data <- read_sav("C:/Users/vasil/Desktop/Korean Conference 2023/Presentations/Abstract 3 TB/FileCollectionTaylor/ARCUN1SF.sav") 
```


## Missing data analysis
```{r echo=TRUE, message=FALSE, warning=FALSE}
c_data <- data %>%
  select(PEQ_Q1_W1,PEQ_Q2_W1,PEQ_Q3_W1,PEQ_Q4_W1,PEQ_Q5_W1,PEQ_Q6_W1)
      

## This calculates % of missing data
Ctotalcells <- prod(dim(c_data))
Cmissingcells <- sum(is.na(c_data))
(Cmissingcells * 100)/(Ctotalcells)

## Testing if missing values are completely at random - MCAR
mcar_test(c_data)

## Highest missing values per variable
c_data %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(), names_to = "names", values_to = "values") %>%
  arrange(desc(values)) %>%
  mutate(perc=(values)/565) %>% # note that 565 is the number of participants and you should replace it with your N 
  slice_head(n=10)
```


## Estimating LPA models (1 to 6) with different number of classes (1:5) 
Here we look for the lowest AIC, BIC, etc. and we follow suggestions outlined in output
```{r echo=TRUE, warning=FALSE, message=FALSE}
set.seed(123)
Models <- c_data%>%
  select("PEQ_Q1_W1","PEQ_Q2_W1","PEQ_Q3_W1","PEQ_Q4_W1","PEQ_Q5_W1","PEQ_Q6_W1")%>%
  single_imputation()%>%
  estimate_profiles(1:5, variances = c("equal", "varying", "equal", "varying"),
                    covariances = c("zero", "zero", "equal", "varying"))%>%
  compare_solutions(statistics = c("AIC","BIC","AWE", "CLC", "KIC"))

Models
```


## Estimating best models 

Model 1 is Equal variances and covariances fixed to 0 (CIP) 
Model 2 is Varying variances and covariances fixed to 0 (CVDP) 
Model 3 is Equal variances and equal covariances (CIUP) 
Model 4 and 5 are not able to fit Mclust 
Model 6 is Varying variances and Varying covariances (CVUP) 


```{r, echo=FALSE}
## Here we run models of interest to create a plot with BIC/AIC and entropy
set.seed(124)
CIUP_5 <- c_data%>%
  select("PEQ_Q1_W1","PEQ_Q2_W1","PEQ_Q3_W1","PEQ_Q4_W1","PEQ_Q5_W1","PEQ_Q6_W1")%>%
  single_imputation()%>%
  estimate_profiles(5, variances="equal",covariances="equal")

set.seed(125)
CIUP_4 <- c_data%>%
  select("PEQ_Q1_W1","PEQ_Q2_W1","PEQ_Q3_W1","PEQ_Q4_W1","PEQ_Q5_W1","PEQ_Q6_W1")%>%
  single_imputation()%>%
  estimate_profiles(4, variances="equal",covariances="equal")

set.seed(126)
CIUP_3 <- c_data%>%
  select("PEQ_Q1_W1","PEQ_Q2_W1","PEQ_Q3_W1","PEQ_Q4_W1","PEQ_Q5_W1","PEQ_Q6_W1")%>%
  single_imputation()%>%
  estimate_profiles(3, variances="equal",covariances="equal")

set.seed(127)
CIUP_2 <- c_data%>%
  select("PEQ_Q1_W1","PEQ_Q2_W1","PEQ_Q3_W1","PEQ_Q4_W1","PEQ_Q5_W1","PEQ_Q6_W1")%>%
  single_imputation()%>%
  estimate_profiles(2, variances="equal",covariances="equal")

set.seed(128)
CIP_2 <- c_data%>%
  select("PEQ_Q1_W1","PEQ_Q2_W1","PEQ_Q3_W1","PEQ_Q4_W1","PEQ_Q5_W1","PEQ_Q6_W1")%>%
  single_imputation()%>%
  estimate_profiles(2, variances="equal",covariances="zero")

set.seed(129)
CIP_3 <- c_data%>%
  select("PEQ_Q1_W1","PEQ_Q2_W1","PEQ_Q3_W1","PEQ_Q4_W1","PEQ_Q5_W1","PEQ_Q6_W1")%>%
  single_imputation()%>%
  estimate_profiles(3, variances="equal",covariances="zero")

set.seed(130)
CIP_4 <- c_data%>%
  select("PEQ_Q1_W1","PEQ_Q2_W1","PEQ_Q3_W1","PEQ_Q4_W1","PEQ_Q5_W1","PEQ_Q6_W1")%>%
  single_imputation()%>%
  estimate_profiles(4, variances="equal",covariances="zero")

set.seed(131)
CIP_5 <- c_data%>%
  select("PEQ_Q1_W1","PEQ_Q2_W1","PEQ_Q3_W1","PEQ_Q4_W1","PEQ_Q5_W1","PEQ_Q6_W1")%>%
  single_imputation()%>%
  estimate_profiles(5, variances="equal",covariances="zero")

fit <- as_tibble(rbind(CIUP_2[["model_3_class_2"]][["fit"]],CIUP_3[["model_3_class_3"]][["fit"]],
      CIUP_4[["model_3_class_4"]][["fit"]],CIUP_5[["model_3_class_5"]][["fit"]],
      CIP_2[["model_1_class_2"]][["fit"]],CIP_3[["model_1_class_3"]][["fit"]],
      CIP_4[["model_1_class_4"]][["fit"]],CIP_5[["model_1_class_5"]][["fit"]])) %>%
  select(Model,Classes,LogLik,AIC,BIC,AWE,Entropy,n_min,BLRT_p)

fit
```


```{r, echo=FALSE}
## select parameters to plot
aic22 <- as.numeric(fit %>% select(AIC)%>% slice(1))
aic23 <- as.numeric(fit %>% select(AIC)%>% slice(2))
aic24 <- as.numeric(fit %>% select(AIC)%>% slice(3))
aic25 <- as.numeric(fit %>% select(AIC)%>% slice(4))
aic62 <- as.numeric(fit %>% select(AIC)%>% slice(5))
aic63 <- as.numeric(fit %>% select(AIC)%>% slice(6))
aic64 <- as.numeric(fit %>% select(AIC)%>% slice(7))
aic65 <- as.numeric(fit %>% select(AIC)%>% slice(8))
bic22 <- as.numeric(fit %>% select(BIC)%>% slice(1))
bic23 <- as.numeric(fit %>% select(BIC)%>% slice(2))
bic24 <- as.numeric(fit %>% select(BIC)%>% slice(3))
bic25 <- as.numeric(fit %>% select(BIC)%>% slice(4))
bic62 <- as.numeric(fit %>% select(BIC)%>% slice(5))
bic63 <- as.numeric(fit %>% select(BIC)%>% slice(6))
bic64 <- as.numeric(fit %>% select(BIC)%>% slice(7))
bic65 <- as.numeric(fit %>% select(BIC)%>% slice(8))
awe22 <- as.numeric(fit %>% select(AWE)%>% slice(1))
awe23 <- as.numeric(fit %>% select(AWE)%>% slice(2))
awe24 <- as.numeric(fit %>% select(AWE)%>% slice(3))
awe25 <- as.numeric(fit %>% select(AWE)%>% slice(4))
awe62 <- as.numeric(fit %>% select(AWE)%>% slice(5))
awe63 <- as.numeric(fit %>% select(AWE)%>% slice(6))
awe64 <- as.numeric(fit %>% select(AWE)%>% slice(7))
awe65 <- as.numeric(fit %>% select(AWE)%>% slice(8))
e22 <- as.numeric(fit %>% select(Entropy)%>% slice(1))
e23 <- as.numeric(fit %>% select(Entropy)%>% slice(2))
e24 <- as.numeric(fit %>% select(Entropy)%>% slice(3))
e25 <- as.numeric(fit %>% select(Entropy)%>% slice(4))
e62 <- as.numeric(fit %>% select(Entropy)%>% slice(5))
e63 <- as.numeric(fit %>% select(Entropy)%>% slice(6))
e64 <- as.numeric(fit %>% select(Entropy)%>% slice(7))
e65 <- as.numeric(fit %>% select(Entropy)%>% slice(8))

## Plot best model
`Number of Profiles` <- as.factor(c(2,3,4,5,6,7,8,9))
Profiles <- recode(`Number of Profiles`,"2"="CIUP 2","3"="CIUP 3","4"="CIUP 4","5"="CIUP 5",
                              "6"="CIP 2","7"="CIP 3","8"="CIP 4","9"="CIP 5")
AIC <- c(aic22,aic23,aic24,aic25,aic62,aic63,aic64,aic65)
BIC <- c(bic22,bic23,bic24,bic25,bic62,bic63,bic64,bic65)
AWE <- c(awe22,awe23,awe24,awe25,awe62,awe63,awe64,awe65)
Entropy <- c(e22,e23,e24,e25,e62,e63,e64,e65)
Model_metrics <- data.frame(Profiles,AIC,BIC,AWE,Entropy)

scl = 35000
figure_2 <- Model_metrics %>%
  ggplot(aes(x = Profiles)) + 
  geom_line(aes(y = AIC, colour = "AIC"), size = 1.0, group=1) +
  geom_line(aes(y = BIC, colour = "BIC"), size = 1.0, group=1) +
  geom_line(aes(y = AWE, colour = "AWE"), size = 1.0, group=1) +
  geom_line(aes(y = Entropy*scl, colour = "Entropy"), size = 1.0,group=1) +
  scale_y_continuous(sec.axis = sec_axis(~. /scl , name = "Entropy")) +
  labs(colour = "Index", y = "AIC / BIC") + 
  theme(legend.position="bottom") + 
  scale_colour_manual(values = c(
    AIC = "royalblue4", BIC = "#E69F00", AWE ="#56B4E9", Entropy ="maroon"))

Model_metrics%>% 
  group_by(Profiles)%>%
  mutate(Total = (sum(AIC,BIC,AWE))/3)%>%
  arrange(Total)

ggsave("Figure_2.tiff", figure_2, width = 10, height = 6, dpi = 300)
```



## CIUP with 3 classes
```{r, echo=TRUE, warning=FALSE, message=FALSE}

write_sav(get_data(CIUP_3), "LPA_with_participant_class.sav")

LPA<-read_sav("LPA_with_participant_class.sav")

LPA_data<-LPA%>%
  mutate(ZProteus_1 = scale(PEQ_Q1_W1)) %>%
   mutate(ZProteus_2 = scale(PEQ_Q2_W1)) %>%
   mutate(ZProteus_3 = scale(PEQ_Q3_W1)) %>%
   mutate(ZProteus_4 = scale(PEQ_Q4_W1)) %>%
   mutate(ZProteus_5 = scale(PEQ_Q5_W1)) %>%
   mutate(ZProteus_6 = scale(PEQ_Q6_W1))

data1 <- data %>%
  add_column(LPA$Class)%>%
  rename(PE_Class= "LPA$Class")%>%
  mutate(ZProteus_1 = scale(PEQ_Q1_W1)) %>%
   mutate(ZProteus_2 = scale(PEQ_Q2_W1)) %>%
   mutate(ZProteus_3 = scale(PEQ_Q3_W1)) %>%
   mutate(ZProteus_4 = scale(PEQ_Q4_W1)) %>%
   mutate(ZProteus_5 = scale(PEQ_Q5_W1)) %>%
   mutate(ZProteus_6 = scale(PEQ_Q6_W1)) 

data1$PE_Class <- as.factor(data1$PE_Class)

data1%>%
  group_by(PE_Class)%>%
  count(PE_Class)%>%
  mutate(Perc = (n/565)*100)

write_sav(data1, "dataC.sav")
```

#CIUP WITH 3 CLASSES
```{r, echo=TRUE, warning=FALSE, message=FALSE}




LPA_data <- LPA_data%>%
  select(Class, ZProteus_1, ZProteus_2, ZProteus_3, ZProteus_4, ZProteus_5, ZProteus_6)%>%
  group_by(Class)%>%
  summarise(ZProteus_1=mean(ZProteus_1),ZProteus_2=mean(ZProteus_2),
            ZProteus_3=mean(ZProteus_3),
            ZProteus_4=mean(ZProteus_4),ZProteus_5=mean(ZProteus_5),ZProteus_6=mean(ZProteus_6))%>%
  na.omit()
LPA_data

LPA_data1 <- LPA_data%>%
  pivot_longer(cols=c(ZProteus_1,ZProteus_2,ZProteus_3,ZProteus_4,ZProteus_5,ZProteus_6),
               names_to = "Model_Indicators", values_to = "Z_Scores")

LPA_data1$Class <- as.factor(LPA_data1$Class)
levels(LPA_data1$Class) <- c("class_1","class_2","class_3")

Figure_3 <- LPA_data1%>%
  ggplot(aes(x = Model_Indicators, y = Z_Scores, group = Class, color=Class)) + 
  geom_point(size = 1.8) + geom_line(size = 1.0) +
  labs(x= "", y = "Z scores", color = "Profiles") + 
  theme(axis.title.x = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1.0),
        axis.title.y = element_text(face = "bold"),
        legend.title = element_text(face = "bold"), 
        legend.position="bottom") +
  scale_y_continuous(breaks=seq(-1.0, 2.0, by = 0.5)) +
  scale_x_discrete(labels=c("ZProteus_1"="Proteus_1","ZProteus_2"="Proteus_2",
                            "ZProteus_3"="Proteus_3", 
                            "ZProteus_4"="Proteus_4",
                            "ZProteus_5"="Proteus_5",
                            "ZProteus_6"="Proteus_6")) +
  scale_colour_manual(values = c(
    Class_1 = "royalblue4", `Class_2` = "#E69F00", `Class_3` ="#56B4E9"))


Figure_3
ggsave("Figure_3.tiff", Figure_3, width = 10, height = 6, dpi = 300)
```

```{r}
Comparisons <- read_sav("C:/Users/vasil/Desktop/Korean Conference 2023/Presentations/Abstract 3 TB/FileCollectionTaylor/Comparisons.sav")
```



```{r}
options(scipen = 999, digits=3, max.print=999999, show.signif.stars=TRUE)
one.way <- aov(IGDTotalW1 ~ PE_Class, data = Comparisons)

summary(one.way)

posthoc_bonferroni <- pairwise.t.test(Comparisons$IGDTotalW1, Comparisons$PE_Class, p.adjust.method = "bonferroni")
print(posthoc_bonferroni)

```
```{r}
options(scipen = 999, digits=3, max.print=999999, show.signif.stars=TRUE)
one.way <- aov(IGDTotalW2 ~ PE_Class, data = Comparisons)

summary(one.way)

posthoc_bonferroni <- pairwise.t.test(Comparisons$IGDTotalW2, Comparisons$PE_Class, p.adjust.method = "bonferroni")
print(posthoc_bonferroni)

```
```{r}
options(scipen = 999, digits=3, max.print=999999, show.signif.stars=TRUE)
one.way <- aov(AverageDailyWeeklySteps~ PE_Class, data = Comparisons)

summary(one.way)

posthoc_bonferroni <- pairwise.t.test(Comparisons$AverageDailyWeeklySteps, Comparisons$PE_Class, p.adjust.method = "bonferroni")
print(posthoc_bonferroni)

```
```{r}
options(scipen = 999, digits=3, max.print=999999, show.signif.stars=TRUE)
one.way <- aov(AverageDailyStepsWave2~ PE_Class, data = Comparisons)

summary(one.way)

posthoc_bonferroni <- pairwise.t.test(Comparisons$AverageDailyStepsWave2, Comparisons$PE_Class, p.adjust.method = "bonferroni")
print(posthoc_bonferroni)

```

```{r}
options(scipen = 999, digits=3, max.print=999999, show.signif.stars=TRUE)
one.way <- aov(AverageDailyStepsWave2~ PE_Class, data = Comparisons)

summary(one.way)

posthoc_bonferroni <- pairwise.t.test(Comparisons$AverageDailyStepsWave2, Comparisons$PE_Class, p.adjust.method = "bonferroni")
print(posthoc_bonferroni)

```

```{r}
options(scipen = 999, digits=3, max.print=999999, show.signif.stars=TRUE)
echo=TRUE
warnings=FALSE
error=FALSE
model1 <- lm( AverageDailyWeeklySteps~ IGDTotalW1, data = Comparisons)
summary(model1)
```
```{r}
options(scipen = 999, digits=3, max.print=999999, show.signif.stars=TRUE)
echo=TRUE
warnings=FALSE
error=FALSE
model2 <- lm( AverageDailyWeeklySteps~ PETotalW1, data = Comparisons)
summary(model2)
```
```{r}
options(scipen = 999, digits=3, max.print=999999, show.signif.stars=TRUE)
echo=TRUE
warnings=FALSE
error=FALSE
model3 <- lm( AverageDailyWeeklySteps~ PETotalW1+IGDTotalW1, data = Comparisons)
summary(model3)
```
```{r}
options(scipen = 999, digits=3, max.print=999999, show.signif.stars=TRUE)
echo=TRUE
warnings=FALSE
error=FALSE
model4 <- lm(AverageDailyCaloriesWave2~ IGDTotalW1, data = Comparisons)
summary(model4)
```
```{r}
options(scipen = 999, digits=3, max.print=999999, show.signif.stars=TRUE)
echo=TRUE
warnings=FALSE
error=FALSE
model5 <- lm(AverageDailyCaloriesWave2~ PETotalW1, data = Comparisons)
summary(model5)
```
```{r}
options(scipen = 999, digits=3, max.print=999999, show.signif.stars=TRUE)
echo=TRUE
warnings=FALSE
error=FALSE
model6 <- lm(AverageDailyCaloriesWave2~ PETotalW1+IGDTotalW1, data = Comparisons)
summary(model6)
```